### 2017-04-22 paper/jfp-submission/safe.tex
some text

We categorized the remaining 42 failures into four groups:
%
\begin{enumerate}
\item 7 infinite type. program doesn't go wrong, but could search for
  context that would go wrong
\item 15 monomorphic functions. recall we abstract a function's type as
  $\tfun$ rather than $\alpha \to \beta$. (may also admit witness by
  searching for context?)
\item 6 dead code. type error occurs in unreachable code, ``deficiency''
  of the type system.
\item 14 bad call. function is well-typed, student calls makes ill-typed
  call, but no error triggered. nothing to be done, as we do not control
  call-site.
\end{enumerate}

>>>>

We categorized the remaining 42 failures into four groups.

\paragraph{Dead Code}
%
Six programs (or 12\% the samples) contained type errors that were
unreachable by any inputs.
%
While technically safe, dead code is generally considered a ``code
smell''; students would likely benefit from a warning in this case.

\paragraph{Bad Function Call}
Fourteen programs contained a function call where the student had
supplied incorrect inputs, yet no error was triggered.
%
Consider, for example the following |assoc| function, which
looks up a key in an \emph{association list} and returns a default if
the key cannot be found.
%
\begin{code}
  let rec assoc (d, k, l) = match l with
    | (ki, vi)::tl ->
       if ki = k then
         vi
       else
         assoc (d, k, tl)
    | _ -> d

  let _ = assoc ([], 123, [(123, "sad"); (321, "happy")])
\end{code}
%
The student's definition of |assoc| is correct, but \ocaml rejects their
subsequent call because the default value |[]| is incompatible with the
|string| values in the list.
%
In this particular call, the key |123| is in the list, so the default
will not be used (and even if it were, no error would be triggered) and
\ocaml's complaint is moot.
%
Of course, \ocaml cannot be expected to know that this particular call
is safe, its type system is not sophisticated enough to express the
necessary conditions.

We note that all of the programs in this category appear to come from
interactions with the \ocaml top-level, \ie the student defined a
function and is now experimenting with it.
%
These calls are thus particularly harmless, indeed they may even be
\emph{beneficial}, as the student may next try looking up a key that
\emph{does not} exist, and realize that the returned values are
incompatible.

% \begin{enumerate}
% \item 7 infinite type. program doesn't go wrong, but could search for
%   context that would go wrong
% \item 15 monomorphic functions. recall we abstract a function's type as
%   $\tfun$ rather than $\alpha \to \beta$. (may also admit witness by
%   searching for context?)
% \item 6 dead code. type error occurs in unreachable code, ``deficiency''
%   of the type system.
% \item 14 bad call. function is well-typed, student calls makes ill-typed
%   call, but no error triggered. nothing to be done, as we do not control
%   call-site.
% \end{enumerate}
